A Case Study on Abstractions to Converge High-Throughput and High-Performance Computing

### 2017-03-24 PandaEvolution.tex
Fix bibliography, the joy of latex
Address Sergey's comments and diagram PanDA architecture

The Titan supercomputer, current number three on the Top 500 list~\cite{top500},
is a Cray XK7 system with 18,688 worker nodes and a total of 299,008 CPU cores.
Each worker node has an AMD Opteron  6274 16-core CPU, a Nvidia Tesla K20X GPU,
32 GB of RAM and no local storage, though a 16 GB RAM disk can be set up. Work
nodes use Cray’s Gemini interconnect for inter-node MPI messaging. Titan is
served by the Spider II~\cite{spider2}, a Lustre filesystem with 32 PB of disk
storage, and by a 29 PB HPSS tape storage system. Titan’s worker nodes run
Compute Node Linux, a run time environment based on SUSE Linux Enterprise
Server.

@@

Currently, no HEP application can benefit from Titan's GPUs but some
computationally-intensive and non memory-intensive tasks can be offloaded from
the Grid to Titan's large amount of cores. Further, when HEP tasks can be
partitioned to independent jobs, Titan worker nodes can be used to execute up to
16 concurrent jobs' payload, one for each available core. Given these
constraints and challenges, the type of task most suitable for execution at the
moment on Titan is Monte Carlo
% event generation and
detector simulation.
% Both types of task are
This type of task is mostly computational-intensive, requiring less than 2GB of
RAM at runtime and with small input data requirements. Detector simulation tasks
in ATLAS account for XX--XX\% of all jobs on WLCG, corresponding to X/X of the
current Grid capacity.\mtnote{Please provide numbers to replace `XX' and,
possibly, a diagram illustrating detector simulation stats compared to the whole
ATLAS computation activity.}

Detector simulation is part of the ATLAS production Monte Carlo (MC) workflow
(also known as MC production
chain)~\cite{rimoldi2006atlas,de2013delphes,ritsch2014atlas}. The MC workflow
consists of four main stages: event generation, detector simulation,
digitization, and reconstruction. Event generation creates sets of particle
four-momenta via different generators, e.g., PYTHIA~\cite{sjostrand2006pythia}
and HERWIG~\cite{corcella2001herwig}. The detector simulator is called
Geant4~\cite{agostinelli2003geant4} and simulates the interaction of these
particles with the sensitive material of the ATLAS detector. Each interaction
creates a so-called hit and all hits are collected and passed on for
digitalization where hits are further process to mimic the readout of the
detector. Finally, reconstruction operates local pattern recognition, creating
high-level objects like particles and jets.

@@

\mtnote{Please feel free to add details about backfill as needed.}
\mtnote{Please note: I know the following departs for the current name given to
the PanDA subsystem installed on Titan. I think there is a good reason to call
it a broker instead of a pilot, and I think I explained it in the previous
paragraphs. Please  take this just as a suggestion, something I would like to
discuss in our meeting.} \sergeynote{It's not a pilot in conventional sense. I
call it an agent. It just happened that we were able to use full PanDA pilot's
code base to serve our purposes on Titan. that's why, by inertia, we still call
it a pilot.}\mtnote{Thank you. We would have a problem calling it `agent' as we
use that term to name the pilot of NGE. Would PanDA Broker work?}\mtnote{Sergey
agreed that we should not use `pilot'. He is quite convinced by `broker' but we
are both open to better suggestions.}

Given the communication requirements of PanDA Pilots and the unused capacity of
Titan, PanDA pilot was redesigned to serve as a job broker on Titan. Architected
as a stand-alone service to be executed on a DTN, this prototype called `PanDA
Broker' implements functionalities to: (i) interrogate Titan about backfill
availability; (ii) pull MC jobs and events; (iii) wrap the payload of ATLAS jobs
into MPI scripts; (iv) submitting MPI scripts to Titan's PBS batch system and
monitor their execution; and (v) staging input/output files. Backfill querying,
payload wrapping, and scripts submission required a new implementation while
pulling ATLAS job and events, and file staging were mostly inherited from PanDA
Pilot.

@@

Once deployed on Titan, every PanDA Broker supports the execution of MC detector
simulation in 8 steps: (i) PanDA Broker queries Titan's Moab scheduler about
current backfill availability; (ii) the PanDA Broker queries the Job Dispatcher
module of the PanDA server for ATLAS jobs that have been bound to Titan by JEDI;
(iii) the PanDA Broker requests to the Data Service module of the PanDA Server
what input data these ATLAS jobs require and transfer the files to the DTN via
the ATLAS DDM service; (iv) the PanDA Broker creates an MPI script, wrapping
enough ATLAS jobs' payload to fit backfill availability; (v) the PanDA Broker
submits the MPI script to the Titan's PBS batch system via RADICAL-SAGA; (vi)
upon execution on the worker node(s), the MPI script creates a ramdisk and
configure AthenaMP to execute the ATLAS jobs' payload; (vii) the payload is
executed into dedicated processes, 16 for each worker node, 1 for each available
core; (viii) upon completion of the payload execution, the PanDA Broker locates
the output on the shared filesystem and transfer it to designated computing
centre.

>>>>

The Titan supercomputer, current number three on the Top 500 list~\cite{top500},
is a Cray XK7 system with 18,688 worker nodes and a total of 299,008 CPU cores.
Each worker node has an AMD Opteron  6274 16-core CPU, a Nvidia Tesla K20X GPU,
32 GB of RAM and no local storage, though a 16 GB RAM disk can be set up. Work
nodes use Cray’s Gemini interconnect for inter-node MPI messaging. Titan is
served by the Spider II~\cite{oral2013olcf}, a Lustre filesystem with 32 PB of
disk storage, and by a 29 PB HPSS tape storage system. Titan’s worker nodes run
Compute Node Linux, a run time environment based on SUSE Linux Enterprise
Server.

@@

Currently, very few HEP applications could benefit from Titan's GPUs but some
computationally-intensive and non memory-intensive tasks of ATLAS workflows can
be offloaded from the Grid to Titan's large amount of cores. Further, when HEP
tasks can be partitioned to independent jobs, Titan worker nodes can be used to
execute up to 16 concurrent payloads, one for each available core. Given these
constraints and challenges, the type of task most suitable for execution at the
moment on Titan is Monte Carlo detector simulation. This type of task is mostly
computational-intensive, requiring less than 2GB of RAM at runtime and with
small input data requirements. Detector simulation tasks in ATLAS account for
$\approx$ 60\% of all the jobs on WLCG, making them a primary candidate for
offloading.

Detector simulation is part of the ATLAS production Monte Carlo (MC) workflow
(also known as MC production
chain)~\cite{rimoldi2006atlas,de2013delphes,ritsch2014atlas}. The MC workflow
consists of four main stages: event generation, detector simulation,
digitization, and reconstruction. Event generation creates sets of particle
four-momenta via different generators, e.g., PYTHIA~\cite{sjostrand2006pythia},
HERWIG~\cite{corcella2001herwig} and many others. The detector simulator is
called Geant4~\cite{agostinelli2003geant4} and simulates the interaction of
these particles with the sensitive material of the ATLAS detector. Each
interaction creates a so-called hit and all hits are collected and passed on for
digitalization where hits are further process to mimic the readout of the
detector. Finally, reconstruction operates local pattern recognition, creating
high-level objects like particles and jets.

@@

\mtnote{Please feel free to add details about backfill as needed.}
\mtnote{Please note: I know the following departs for the current name given to
the PanDA subsystem installed on Titan. I think there is a good reason to call
it a broker instead of a pilot, and I think I explained it in the previous
paragraphs. Please  take this just as a suggestion, something I would like to
discuss in our meeting.} \sergeynote{It's not a pilot in conventional sense. I
call it an agent. It just happened that we were able to use full PanDA pilot's
code base to serve our purposes on Titan. that's why, by inertia, we still call
it a pilot.}\mtnote{Thank you. We would have a problem calling it `agent' as we
use that term to name the pilot of NGE. Would PanDA Broker work?}\sergeynote{OK
broker it is for this paper.}

Given the communication requirements of PanDA Pilots and the unused capacity of
Titan, PanDA pilot was repurposed to serve as a job broker on the DTN nodes of
Titan. Maintaining the core modules of PanDA Pilot and its stand-alone
architecture, this prototype called `PanDA Broker' implements functionalities
to: (i) interrogate Titan about backfill availability; (ii) pull MC jobs and
events; (iii) wrap the payload of ATLAS jobs into MPI scripts; (iv) submitting
MPI scripts to Titan's PBS batch system and monitor their execution; and (v)
staging input/output files. Backfill querying, payload wrapping, and scripts
submission required a new implementation while pulling ATLAS job and events, and
file staging were inherited from PanDA Pilot.

@@

Once deployed on Titan, every PanDA Broker supports the execution of MC detector
simulation in 8 steps: (i) PanDA Broker queries Titan's Moab scheduler about
current backfill availability; (ii) the PanDA Broker queries the Job Dispatcher
module of the PanDA server for ATLAS jobs that have been bound to Titan by JEDI;
(iii) Upon receiving the descriptions of those jobs, PanDA Broker pulls their
input files from the ATLAS DDM service to the DTN; (iv) the PanDA Broker creates
an MPI script, wrapping enough ATLAS jobs' payload to fit backfill availability;
(v) the PanDA Broker submits the MPI script to the Titan's PBS batch system via
RADICAL-SAGA; (vi) upon execution on the worker node(s), the MPI script creates
configures and executes one AthenaMP for each work node available; (vi) AthenaMP
spawns 1 event simulation process on each available core (16); (vii) during
execution, PanDA Broker monitors execution progress and sends PanDA Server
``heart beats'' for each job.; and (viii) upon completion of each simulation,
the PanDA Broker locates the output on the shared filesystem, transfer it to
designated computing centre, and performs cleanup.

### 2017-03-24 PandaOverview.tex
Address Sergey's comments and diagram PanDA architecture

\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figures/PandaArch.jpg}
  \end{center}
\caption{\mtnote{This diagram is a placeholder. I am working on a substitute
that will be referenced when describing PanDA's subsystems.}Schematic view of
the PanDA system, as deployed for production in 2017. Black boxes indicate
PanDA's subsystems; Orange boxes components of a subsystem; purple boxes
databases; green arrows http messaging interfaces; red arrows API; purple arrows
database communications.}
% Originally PanDA was designed for grid infrastructure... In this
% paper is focussed on the HPC use case; there are differences from
% the grid use case. This paper discusses how PanDA has been adapted
% to execute a workload % on HPC resources. The schematic overview is
% presented for workload X on Titan......
\label{fig:architecture}
\end{figure}

@@

Upon bootstrapping, PanDA Pilot downloads major parts of its runtime code via
HTTP from a central Subversion repository. The repository works with an Apache
web server, configured with a memory-based web proxy (Squid). The purpose of the
cache is to reduce the request load on the back-end Subversion server. That
improves performance \mtnote{Do we have data/numbers?} since only occasional
queries trigger a full lookup on the back-end Subversion system, and most
external queries are pulled from memory on the front-end web server. The
benefits of this system include a high-performance code download service,
combined with code updates still being immediately available as soon as they are
committed to source code control.

>>>>

\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figures/panda_architecture.pdf}
  \end{center}
\caption{PanDA architecture representing the subsystems and components relevant
to the integration of PanDA with Titan. Numbers indicates the execution process
based on JEDI, from submission of tasks (1) to retrieval of their output (20).
The monitoring subsystem, the architectural details of PanDA Pilot and the
communication among subsystem's components are abstracted to improve clarity.
% \mtnote{This diagram is a placeholder. I am working on a substitute
% that will be referenced when describing PanDA's subsystems.}Schematic view of
% the PanDA system, as deployed for production in 2017. Black boxes indicate
% PanDA's subsystems; Orange boxes components of a subsystem; purple boxes
% databases; green arrows http messaging interfaces; red arrows API; purple
% arrows database communications.
}
% Originally PanDA was designed for grid infrastructure... In this
% paper is focussed on the HPC use case; there are differences from
% the grid use case. This paper discusses how PanDA has been adapted
% to execute a workload % on HPC resources. The schematic overview is
% presented for workload X on Titan......
\label{fig:architecture}
\end{figure}

@@

Upon bootstrapping, PanDA Pilot downloads major parts of its runtime code via
HTTP from a central Subversion repository. The repository works with an Apache
web server, configured with a memory-based web proxy (Squid). The purpose of the
cache is to reduce the request load on the back-end Subversion server. That
improves performance since only occasional queries trigger a full lookup on the
back-end Subversion system, and most external queries are pulled from memory on
the front-end web server. The benefits of this system include a high-performance
code download service, combined with code updates still being immediately
available as soon as they are committed to source code control.
